Should I AirBNB My Property?

### 2016-12-12 abbReturns_Script.Rmd
REORG:  Through final

This document is intended the literate programming of our analysis for the 'Should I AirBNB My Property?' research project.  Below we document the complete data provenance, taking our data from the raw source data through to the final results. Data prepartion, modeling and visualization will be done in R (version 3.3.1). As a very brief overview, we will be comparing the financial returns of short-term (AirBNB-type) and long-term (traditional leases) leasing of residential properties in Melbourne, Australia.  We perform this comparison under a hypothetical condition where-in an investor purchases a property on Septemeber 1, 2015 and is looking to determine whether too lease out the property on a traditional long-term lease (12 months) or as an AirBNB property.  We examine the returns for the following 12 months.  Sensitivity to time frames, start dates, locations, property types, etc. are considered in this analysis as well.

Data on AirBNB (short-term) ltrals has been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company.

Throughout this document we will refer to the short-term data as **'short-term'** or **'AirBNB'** while the long-term data will be referred to as **'long-term'**, **'ltral'** or **'lease'**.  Short-term nightly fees will be referenced as **'rates'** and long-term as **'ltr'**.  We will use **'prices'** to collectively refer to **rates** and **ltrs**. Note that within the Melbourne market ltrals are quoted in weekly ltrs (but are charged to tenants at a monthly ltr equivalent to the 1/12 of 52 weeks of ltr).

@@

 ## Set path location

  # Get computer names
  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name
  if(comp.name == '7020D-121777-W' | 
      comp.name == 'DESKTOP-1D7JO4J'){

    data.path <- 'c:/dropbox/research/airBNB/data/'

@@

    data.path <- 'gideon path'

@@

 ## Load sources (custom function files)

  source("c:/code/dataviztools/ggplottools.R")
  source("c:/code/datamgmttools/dataMungetools.R")
  source("c:/code/research/AirBNBMelbourne/analysis_Functions.R")
  source("c:/code/research/AirBNBMelbourne/dataPrep_Functions.R")

 ## Set a global parameter

@@

@@

 ## Set basic color to use in palettes

@@

@@

Both the AirBNB and the ltral data require some measure of data pre-processing and due to the run times of these function they are done outside of this document.  For the AirBNB data, the pre-processing is primarily concerned with imputing missing daily observations for a small number of properties in the dataset.  Full information on this process can be found in the **airbnbDataClean.Rmd** and **dataCleanFunctions.R** files.  Do note that most of the imputed data is ignored by our choice of study time period later on.

@@

Next we load in the raw data.  The short-term data is found in two separate files, one providing information on the property itself data and one providing a daily record of bookings.

```{r load_abb_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

@@

We then load the long term ltral data.  This includes property level data (including the final lease rate) and listing-level data containing information on all changes to listing prices and dates, etc.

@@

 ## Property level long term ltral information

@@

 ## Long term ltral listing information

@@

And, finally, we load two GIS shapefiles that denote suburb and Statistical Area 1 (sa1) boundaries in the Melbourne area. We also load a shapefile that contains the location of the beach areas along Port Phillip Bay.^[Only those east of the mouth of Yarra River.]

@@

 ## SA1 shapefile

@@

  ## Suburb shapefile

  ## Beach file

@@

## Data Cleaning and Preparation

@@

### Trim Fields

@@

@@

 ## Long term data

@@

 ## Geographic data

@@

 ## Property Data

@@

  abb.data <- abb.data[ , abb.keep.flds]

@@

 ## Long term property Data

@@

  ltr.data <- ltr.data[ , ltr.keep.flds]

@@

### Standardize Field Names

Next we standardize the field names.  Here we follow the short-term data convention of placing a period between words in the name.  We also shorten a number of the fields to make future coding easier.  No changes are made to the monthly or daily data.

For the short term data, it already uses the *word-.-word* convention.  Here our renaming is done to shorten certain field names.

@@

### Field Conversions

@@

Below we convert and coalate a number of the temporal and spatial variables.

@@

We convert time variables across all dataset to an R date format.  For the AirBNB data this is a relatively straighforword process as the data are already standardized into a consistant date representation. We also create a year variable.

@@

 ## Property Data Fields

@@

  ## Daily Data

@@

The dates in the long term data are not standardized.  To fix this issue we create a small function that will remove the unnecessary digits and spaces and then convert to a standard date form.  We then apply this formula to the three date fields in the long term data.  We also create a year field for the event date.

@@

 ## Build custom function

@@

 ## Convert event date

  # Date
  ltr.data$event.date <- fixLTDates(ltr.data$event.date)

  # Year
  ltr.data$event.year <- as.numeric(substr(ltr.data$event.date, 1, 4))

 ## Converted list data date fields

@@

#### Fix Boolean Variables

Next we fix a few of the boolean variables in the long term data that are stored in a non-compatible format.

@@

AirBNB units are classified into 19 diffeltr property type.  The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that the Igloo, at the very least, is not a completely true representation of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

@@

Within the long-term ltral data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories:

@@

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the AirBNB data conforms to North American lexicon.  This is most appaltr when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).  Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

@@

 ## Short Term data

@@

 ## Long Term data

@@

The **Other** categories contain very small percentage of properties in both the AirBNB and the ltral data.  As a result, we remove the 'Others' from the data.

@@

Next, for the short term ltrals units can be listed as one of three types relating to the extent of the property which is able to be booked:

1. **Entire Home/Apt**: The entire home or apartment is available
2. **Private Room**: One room within a house or apartment is available
3. **Shared Room**: A bed within a room shared by other occupants is available

As long-term ltrals do not offer **Private Room** or **Shared Room** options and our purpose here is a comparison of the two we remove all short term properties that do not lease the entire home/apt. Unfortunately, this filter does remove about 40% of the short term data, however, given the research question it is unavoidable.

@@

@@

  ggplot(abb.data, aes(x=listing.type)) + 
    geom_bar(color=abb.col[1], fill=abb.col[1]) +
    xlab('\nListing Type\n') + 
    ggtitle('\nShort Term ltrals\n')  +
    ylab('# of Properties\n')

@@

### Trim by Time

Next we trim the observations by time.  Our hypothetical simulated decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long term ltral versus an AirBNB, short term ltral approach to generating income.  To determine the likely long term ltral rate for any given property, we will use the observed ltral transactions from the market from September 1, 2014 to August 31, 2015.  Taking the AirBNB approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factor).  In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.  Another reason for using these two period of data is that the AirBNB data has missing daily values prior to September 1, 2016.  While we have imputed these missing daily observations, we hold the actual observed values to be of greater accuracy and have designed the initial simulated example to take advantage of the full observed set.

 ## Limit short term data

  abb.dataf <- abb.data[abb.data$min.date >= '2015-09-01', ]
  daily.dataf <- daily.data[daily.data$date >= '2015-09-01', ]

 ## Limit long term data

  ltr.dataf <- ltr.data[ltr.data$event.date >= '2014-09-01' & 
                            ltr.data$event.date <= '2015-08-31', ]
  listing.dataf <- listing.data[listing.data$id.key %in% ltr.dataf$id.key, ]

@@

Location is clearly an important determinant in both the short and long term ltral markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs.  In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts.  The specific suburb of a property is the second, or finer scale, at which the market operates. The plot below shows the three large submarkets.

@@

 ## Convert shapefile to ggplot-able object

@@

  # Limit to extent of data
  sub.table <- table(ltr.dataf$suburb)
  sub.table <- sub.table[sub.table != 0]

@@

```{r mappoints2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.height=5}

@@

  # Set colors
  g.cols <- abb.col[c(2,10,9)]

@@

## Plot map

@@

It is likely that the short-term ltral market follows a similar market hierarchy, however, we can imagine that a few diffeltr spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beach; and 2) Proximity to key tourist activities and events.  Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analysing short and long term ltrals in Melboure:

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short and long term data.  We also add the Statistical Area 1 code for each property for use in additional spatial modeling later on. Fainlly, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs).

@@

@@

 ## Determine centroid of study area

  latMid <- median(abb.dataf$latitude)
  longMid <- median(abb.dataf$longitude)

 ## Calculate Degress to Meters conversion

@@

 ## Create the beach buffer

@@

@@

  # Convert to point shapefile
  abb.prop.shp <- SpatialPointsDataFrame(cbind(abb.dataf$longitude,
                                               abb.dataf$latitude),
                                         abb.dataf,
                                         proj4string=CRS("+init=epsg:4283"))

  # Add suburb name 
  spJoin <- over(abb.prop.shp, suburbs.shp)
  abb.dataf$suburb <- as.character(spJoin$name_2006)
  abb.dataf$suburb[is.na(abb.dataf$suburb)] <- 'missing'

  # Add suburb designation
  abb.dataf$sub.class <- as.character(spJoin$suburbdesi)
  abb.dataf$sub.class[is.na(abb.dataf$sub.class)] <- 'missing'

  # Add SA1 designation
  spJoin <- over(abb.prop.shp, sa1s.shp)
  abb.dataf$sa1 <- spJoin$sa1_main11

  ## Add Beach Designation
  spJoin <- over(abb.prop.shp, beach.buffer)
  abb.dataf$beach <- ifelse(is.na(spJoin), 0 ,1)

  # Remove those missing a suburb classification (out of area)
  abb.dataf <- abb.dataf[abb.dataf$sub.class != 'missing', ]
  abb.dataf <- abb.dataf[abb.dataf$sa1 != 'missing', ]

@@

  # Add individual classes
  abb.dataf$sub.mrkt <- 'rural'
  abb.dataf$sub.mrkt[abb.dataf$sub.class == 'Middle'] <- 'suburban'
  abb.dataf$sub.mrkt[abb.dataf$sub.class == 'Inner'] <- 'city'
  abb.dataf$sub.mrkt[abb.dataf$suburb %in% core.suburbs] <- 'city-core'
  abb.dataf$sub.mrkt[abb.dataf$beach == 1] <- 'beach'

  # Order the factors properly
  abb.dataf$sub.mrkt <- factor(abb.dataf$sub.mrkt, 
                               levels=c('city-core', 'city', 'suburban', 
                                        'rural', 'beach'))

@@

  # Make map
  abb.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group), color='gray50') +
    geom_point(data=abb.dataf, aes(x=longitude, y=latitude, color=sub.mrkt), size=.2) +
    scale_color_manual(values=sm.col) +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3))) + 
    ggtitle('AirBNB Location\nBy Submarket')

@@

About 5% of the long term ltral observations have missing latitude and longitude values.  Half of these do, however, have a lat/long value for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps for those properties that lack a specific lat/long value we apply the street centroid value.  For those without either (3% of the total) we remove them from the dataset at this point.  After removing these, we apply the submarket designations.

 ## Label those missing lat/long

  miss.LL <- which(is.na(ltr.dataf$latitude))

 ## Create a field to indicate the approximate matches

  ltr.dataf$miss.LL <- 0
  ltr.dataf$miss.LL[miss.LL] <- 1

 ## Apply approximate match

  ltr.dataf$latitude[miss.LL] <- ltr.dataf$street.latitude[miss.LL]
  ltr.dataf$longitude[miss.LL] <- ltr.dataf$street.longitude[miss.LL]

 ## Remove those still missing latitude data

  ltr.dataf <- ltr.dataf[!is.na(ltr.dataf$latitude), ]

@@

 ## Long term data

  # Convert to point shapefile
  ltr.prop.shp <- SpatialPointsDataFrame(cbind(ltr.dataf$longitude,
                                                ltr.dataf$latitude),
                                         ltr.dataf,
                                         proj4string=CRS("+init=epsg:4283"))

  # Add suburb name 
  spJoin <- over(ltr.prop.shp, suburbs.shp)
  ltr.dataf$suburb <- as.character(spJoin$name_2006)
  ltr.dataf$suburb[is.na(ltr.dataf$suburb)] <- 'missing'

  # Add suburb designation
  ltr.dataf$sub.class <- as.character(spJoin$suburbdesi)
  ltr.dataf$sub.class[is.na(ltr.dataf$sub.class)] <- 'missing'

  # Add SA1 designation
  spJoin <- over(ltr.prop.shp, sa1s.shp)
  ltr.dataf$sa1 <- spJoin$sa1_main11

  ## Add Beach Designation
  spJoin <- over(ltr.prop.shp, beach.buffer)
  ltr.dataf$beach <- ifelse(is.na(spJoin), 0 ,1)

  # Remove those missing a suburb classification (out of area)
  ltr.dataf <- ltr.dataf[ltr.dataf$sub.class != 'missing', ]
  ltr.dataf <- ltr.dataf[ltr.dataf$sa1 != 'missing', ]

@@

  # Add individual classes
  ltr.dataf$sub.mrkt <- 'rural'
  ltr.dataf$sub.mrkt[ltr.dataf$sub.class == 'Middle'] <- 'suburban'
  ltr.dataf$sub.mrkt[ltr.dataf$sub.class == 'Inner'] <- 'city'
  ltr.dataf$sub.mrkt[ltr.dataf$suburb %in% core.suburbs] <- 'city-core'
  ltr.dataf$sub.mrkt[ltr.dataf$beach == 1] <- 'beach'

  # Order the factors properly
  ltr.dataf$sub.mrkt <- factor(ltr.dataf$sub.mrkt, 
                               levels=c('city-core', 'city', 'suburban', 
                                        'rural', 'beach'))

@@

  # Make map
  ltr.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group), color='gray50') +
    geom_point(data=ltr.dataf, aes(x=longitude, y=latitude, color=sub.mrkt), size=.1) +
    scale_color_manual(values=sm.col) +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3))) + 
    ggtitle('Long Term ltral Location\nBy Submarket')

@@

@@

  # Make map
  abbtype.map <- ggplot() +
    geom_bar(data=abb.dataf, aes(x=sub.mrkt, fill=sub.mrkt)) +
    scale_fill_manual(values=sm.col) +
    facet_wrap(~type) +
    ylab('# of Properties\n') + xlab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    ggtitle('AirBNB Property Type\nBy Submarket')

>>>>

This document is the literate programming of our analysis for the 'Should I AirBNB My Property?' research project.  Below we document the complete data provenance, taking our data from the raw source data through to the final results. Data prepartion, modeling and visualization will be done in R (version 3.3.1). As a very brief overview, we will be comparing the financial returns of short-term (AirBNB-type) and long-term (traditional leases) leasing of residential properties in Melbourne, Australia.  We perform this comparison under a hypothetical condition where-in an investor purchases a property on Septemeber 1, 2015 and is looking to determine whether to lease out the property on a traditional long-term lease (12 months) or nightly as an AirBNB property.  We examine the returns for the following 12 months.  Sensitivity to time frames, start dates, locations, property types, etc. are considered in this analysis as well.

Data on AirBNB (short-term) rentals have been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company.

Throughout this document we will refer to the short-term data as **'short-term'** or **'AirBNB'** while the long-term data will be referred to as **'long-term'**, **'rental'** or **'lease'**.  Short-term nightly fees will be referenced as **'rates'** and long-term as **'rent'**.  We will use **'prices'** to collectively refer to **rates** and **rents**. Note that within the Melbourne long term rentals are quoted in weekly rents but are charged to tenants at a monthly rent equivalent to 1/12 of 52 weeks of rent.

@@

We then set the paths to the data and the code, conditional on which of the authors is running the analysis.

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  # Get computer names

  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name

  if(comp.name == '7020D-121777-W' | comp.name == 'DESKTOP-1D7JO4J'){

    data.path <- 'c:/dropbox/research/airBNB/data/'
    code.path <- 'c:/code/'

@@

    data.path <- 'alternate data path'
    code.path <- 'alternate data path'

@@

Next, we source the necessary custom code files for the analysis.

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  source(paste0(code.path, "dataviztools/ggplottools.R"))
  source(paste0(code.path, "datamgmttools/dataMungetools.R"))
  source(paste0(code.path, "research/AirBNBMelbourne/analysis_Functions.R"))
  source(paste0(code.path, "research/AirBNBMelbourne/dataPrep_Functions.R"))

We then set the constants used in this study:

1. **Exchange rate**: The Airbnb rates are provided in US dollars.  We convert these to Australian dollars based on the average exchange rate over the Sep 1 2015 to Aug 30 2015 period, or about 1.32 Australian dollars to 1 US dollar.

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

@@

Finally, we set the color to be used in the analysis.  We use AirBnb's official palette of 10 colors.

@@

 # Set all colors

@@

  # Set specific colors

  str.col <- abb.col[1]
  ltr.col <- abb.col[3]

@@

Both the AirBNB and the long term rental data require some measure of data pre-processing and due to the run times of these processes they are done outside of this document.  For the AirBNB data, the pre-processing is primarily concerned with imputing missing daily observations for a small number of properties in the dataset.  Full information on this process can be found in the **airbnbDataClean.Rmd** and **dataCleanFunctions.R** files.  Do note that most of the imputed data is ignored by our choice of study time period.

@@

Next we load in the raw data.  The Airbnb data is found in two separate files, one providing information on the property itself and one providing a daily record of bookings.  
```{r load_abb_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

@@

We then load the long term rental data.  This includes property level data (including the final lease rate) and listing-level data containing information on all changes to listing prices and dates, etc.

@@

 # Property level long term ltral information

@@

 # Long term ltral listing information

@@

And, finally, we load three GIS shapefiles.  The first two denote suburb and Statistical Area 1 (sa1) boundaries in the Melbourne area. We also load a shapefile that contains the location of the beach areas along Port Phillip Bay.^[Only those east of the mouth of Yarra River as those to the west are primarily industrial land.]

@@

  # Suburb shapefile

  # SA1 shapefile

@@

  # Beach file

@@

## Data Preparation

@@

### Trim and Standardize Fields

@@

 # Short term data

@@

 # Long term data

@@

 # Geographic data

@@

@@

  abb.data <- abb.data[, abb.keep.flds]

@@

@@

  ltr.data <- ltr.data[,ltr.keep.flds]

@@

Next we standardize the field names.  Here we follow the short-term data convention of placing a period between words in the name.  We also shorten a number of the fields to make future coding easier.

For the short term data, it already uses the *word-.-word* convention.  Here our renaming is done to shorten certain field names. No changes are made to the names of the Airbnb daily data.

@@

### Temporal Considerations

@@

Below we rename, transform and filter our data to fit the time period appropriate to our research question.

@@

We convert time variables across all dataset to an R date format.  For the AirBNB data this is a relatively straighforword process as the data are already standardized into a consistant date representation. We fix dates in both the property and the daily data.

@@

  # Property Data Fields

@@

  # Daily Data

@@

The dates in the long term data are not standardized.  To fix this issue we create a small function that will remove the unnecessary digits and spaces and then convert to a standard date form.  We then apply this formula to the three date fields in the long term data (both listing and rental).  We also create a year field for the event date.

@@

 # Build custom function

@@

  # Convert event date

  ltr.data$event.date <- fixLTDates(ltr.data$event.date)

  # Extract Year of event date
  ltr.data$event.year <- as.numeric(substr(ltr.data$event.date, 1, 4))

  # Converted listing data date fields

@@

#### Trim by Time

Next we trim the observations by time.  Our hypothetical simulated decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long term rental versus an AirBNB, short term rental approach to generating income.  To determine the likely long term rental rate for any given property, we will use the observed rental transactions from the market from September 1, 2014 to August 31, 2015.  Taking the AirBNB approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors).  In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.  Another reason for using these two period of data is that the AirBNB data has missing daily values prior to September 1, 2015.  While we have imputed these missing daily observations, we hold the actual observed values to be of greater accuracy and have designed the initial simulated example to take advantage of the full observed set.

 # Limit short term data

  abb.data <- abb.data[abb.data$min.date >= '2015-09-01', ]
  daily.data <- daily.data[daily.data$date >= '2015-09-01', ]

 # Limit long term data

  ltr.data <- ltr.data[ltr.data$event.date >= '2014-09-01' & 
                            ltr.data$event.date <= '2015-08-31', ]
  listing.data <- listing.data[listing.data$id.key %in% ltr.data$id.key, ]

@@

AirBNB units are classified into 19 different property type.  The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that the Igloo, at the very least, is not a completely true representation of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

@@

Within the long-term rental data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories:

@@

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the AirBNB data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).  Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

@@

 # Short Term data

@@

 # Long Term data

@@

As the chart below show, the **Other** category only makes up a very small percentage of the properties in both the short and long term market, especially in the long-term market.  Additionally, a good portion of the **Others** in the Airbnb data are Bed & Breakfast units, a use which doesn't fit with our research question.  As a result, we remove the 'Others' from both the Airbnb and the long term rental data.

```{r type_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE, fig.height=4}

  # Create Airbnb plot

  abb.types <- ggplot(abb.data, aes(x=type)) + 
    geom_bar(color=str.col, fill=str.col) +
    xlab('\nProperty Type\n') + 
    ggtitle('\nShort Term Rentals\n')  +
    ylab('# of Properties\n')

  # Create long term rental plot

  ltr.types <- ggplot(ltr.data, aes(x=type)) + 
    geom_bar(color=ltr.col, fill=ltr.col) +
    xlab('\nProperty Type\n') + 
    ggtitle('\nLong Term Rentals\n')  +
    ylab('# of Properties\n')

  # Output plots

  ggMultiPlots(abb.types, ltr.types, cols=2)

@@

Next, for the short term rentals properties can be listed as one of three types relating to the extent of the property which is able to be booked:

1. **Entire Home/Apt**: The entire home or apartment is available
2. **Private Room**: One room within a house or apartment is available
3. **Shared Room**: A bed within a room shared by another occupant(s) is available

As long-term rentals do not offer **Private Room** or **Shared Room** options^[Our dataset does not include rooming houses, purpose built student accomodations and other types of rental which may offer private or shared room options] and our purpose here is a comparison of long and short term returnw, we remove all short term properties that do not lease the entire home/apt. Unfortunately, this filter does remove about 40% of the short term data, however, given the research question it is unavoidable.

@@

  # Fix data types and names

@@

  # Make Plot

  ggplot(abb.data, aes(x=listing.type)) + 
    geom_bar(color=abb.col[1], fill=abb.col[1]) +
    xlab('\nListing Type\n') + 
    ggtitle('\nShort Term Rentals by Listing Type\n')  +
    ylab('# of Properties\n')

@@

@@

Location is clearly an important determinant in both the short and long term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs.  In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts.  The specific suburb of a property is the second, or finer scale, at which the market operates. The plot below shows the three large submarkets (Inner, Middle and Outer).

@@

 # Convert shapefile to ggplot-able object

@@

  # Limit to extent of data
  sub.table <- table(ltr.data$suburb)
  sub.table <- sub.table[sub.table != 0]

@@

```{r mappoints2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

@@

  # Set colors
  g.cols <- abb.col[c(2, 10, 9)]

@@

```{r plot_loc_map, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.height=6}

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beach; and 2) Proximity to key tourist activities and events.  Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analysing short and long term ltrals in Melboure:

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short and long term data.  We also add the Statistical Area 1 code for each property for use in additional spatial modeling later on. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs).

We select the 18 suburbs below to represent the 'core' of the city.  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well serviced by public transportation and possess abundant amenties for tourists.

@@

We then apply a 500 meter buffer to the beach along Port Phillip Bay.

@@

 # Determine centroid of study area

  latMid <- median(abb.data$latitude)
  longMid <- median(abb.data$longitude)

 # Calculate Degress to Meters conversion

@@

 # Create the beach buffer

@@

Next we assign the suburb, SA1, city-core and beach designations to observations in the Airbnb dataset.

@@

  # Convert to point shapefile
  abb.prop.shp <- SpatialPointsDataFrame(cbind(abb.data$longitude,
                                               abb.data$latitude),
                                         abb.data,
                                         proj4string=CRS("+init=epsg:4283"))

  # Add suburb name 
  spJoin <- over(abb.prop.shp, suburbs.shp)
  abb.data$suburb <- as.character(spJoin$name_2006)
  abb.data$suburb[is.na(abb.data$suburb)] <- 'missing'

  # Add suburb designation
  abb.data$sub.class <- as.character(spJoin$suburbdesi)
  abb.data$sub.class[is.na(abb.data$sub.class)] <- 'missing'

  # Add SA1 designation
  spJoin <- over(abb.prop.shp, sa1s.shp)
  abb.data$sa1 <- spJoin$sa1_main11

  ## Add Beach Designation
  spJoin <- over(abb.prop.shp, beach.buffer)
  abb.data$beach <- ifelse(is.na(spJoin), 0 ,1)

  # Remove those missing a suburb classification (out of area)
  abb.data <- abb.data[abb.data$sub.class != 'missing', ]
  abb.data <- abb.data[abb.data$sa1 != 'missing', ]

@@

  # Add individual classes
  abb.data$sub.mrkt <- 'rural'
  abb.data$sub.mrkt[abb.data$sub.class == 'Middle'] <- 'suburban'
  abb.data$sub.mrkt[abb.data$sub.class == 'Inner'] <- 'city'
  abb.data$sub.mrkt[abb.data$suburb %in% core.suburbs] <- 'city-core'
  abb.data$sub.mrkt[abb.data$beach == 1] <- 'beach'

  # Order the factors properly
  abb.data$sub.mrkt <- factor(abb.data$sub.mrkt, 
                              levels=c('city-core', 'city', 'suburban', 
                                        'rural', 'beach'))

@@

Before we do the same for the long term data, we must remedy the fact that about 5% of the long term rental observations have missing latitude and longitude values.  Half of these do, however, have a lat/long value for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps for those properties that lack a specific lat/long value we apply the street centroid value.  For those without either (3% of the total) we remove them from the dataset at this point.

 # Label those missing lat/long

  miss.LL <- which(is.na(ltr.data$latitude))

 # Create a field to indicate the approximate matches

  ltr.data$miss.LL <- 0
  ltr.data$miss.LL[miss.LL] <- 1

 # Apply approximate match

  ltr.data$latitude[miss.LL] <- ltr.data$street.latitude[miss.LL]
  ltr.data$longitude[miss.LL] <- ltr.data$street.longitude[miss.LL]

 # Remove those still missing latitude data

  ltr.data <- ltr.data[!is.na(ltr.data$latitude), ]

@@

Having fixed the missing lat/long issue, we then assign the suburb, SA1, city-core and beach designations to observations in the long term rental dataset.

@@

  # Long term data

  # Convert to point shapefile

  ltr.prop.shp <- SpatialPointsDataFrame(cbind(ltr.data$longitude,
                                               ltr.data$latitude),
                                         ltr.data,
                                         proj4string=CRS("+init=epsg:4283"))

  # Add suburb name

  spJoin <- over(ltr.prop.shp, suburbs.shp)
  ltr.data$suburb <- as.character(spJoin$name_2006)
  ltr.data$suburb[is.na(ltr.data$suburb)] <- 'missing'

  # Add suburb designation

  ltr.data$sub.class <- as.character(spJoin$suburbdesi)
  ltr.data$sub.class[is.na(ltr.data$sub.class)] <- 'missing'

  # Add SA1 designation

  spJoin <- over(ltr.prop.shp, sa1s.shp)
  ltr.data$sa1 <- spJoin$sa1_main11

  ## Add Beach Designation

  spJoin <- over(rent.prop.shp, beach.buffer)
  ltr.data$beach <- ifelse(is.na(spJoin), 0 ,1)

  # Remove those missing a suburb classification (out of area)

  ltr.data <- ltr.data[ltr.data$sub.class != 'missing', ]
  ltr.data <- ltr.data[ltr.data$sa1 != 'missing', ]

@@

  # Add individual classes
  ltr.data$sub.mrkt <- 'rural'
  ltr.data$sub.mrkt[ltr.data$sub.class == 'Middle'] <- 'suburban'
  ltr.data$sub.mrkt[ltr.data$sub.class == 'Inner'] <- 'city'
  ltr.data$sub.mrkt[ltr.data$suburb %in% core.suburbs] <- 'city-core'
  ltr.data$sub.mrkt[ltr.data$beach == 1] <- 'beach'

  # Order the factors properly
  ltr.data$sub.mrkt <- factor(ltr.data$sub.mrkt, 
                              levels=c('city-core', 'city', 'suburban', 
                                        'rural', 'beach'))

The location of the short and long term rentals, colored by submarket, are shown below

  # Set submarket colors

  # Make map

  abb.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group), color='gray50') +
    geom_point(data=abb.data, aes(x=longitude, y=latitude, color=sub.mrkt), size=.2,
               alpha=.2) +
    scale_color_manual(values=sm.col) +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3,
                                                     alpha=1))) + 
    ggtitle('AirBNB Location\nBy Submarket')

@@

  # Make map

  ltr.map <- ggplot() +
    geom_polygon(data=subs, aes(x=long, y=lat, group=group), color='gray50') +
    geom_point(data=ltr.data, aes(x=longitude, y=latitude, color=sub.mrkt), size=.1,
               alpha=.1) +
    scale_color_manual(values=sm.col) +
    scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
    scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
    xlab('') + ylab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3,
                                                     alpha=1))) + 
    ggtitle('Long Term Rental Location\nBy Submarket')

@@

The chart below break down the relative frequency of short and long term rental properties across the five submarkets.

@@

  # Make Airbnb submarket map

  abbtype.map <- ggplot() +
    geom_bar(data=abb.dataf, aes(x=sub.mrkt, fill=sub.mrkt)) +
    scale_fill_manual(values=sm.col) +
    facet_wrap(~type) +
    ylab('# of Properties\n') + xlab('')+
    theme(legend.position = 'bottom',
          legend.title = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    ggtitle('AirBNB Property Type\nBy Submarket')

  # Make Long term rental map
